### 2013-06-07 drafts/general_writeup/writeup.tex
Started section on learning physics.

\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{amsmath} 
\usepackage{algpseudocode}
\usepackage{ amssymb }
\usepackage{listings}
\usepackage{syntax}
\usepackage{mdframed}

@@

Our probabilistic model is motivated by the fact that while $\mathcal{G}$
represents the language in which solutions are written, in any particular
domain there is a distribution $\mathcal{H}$ over solutions that is itself drawn
according to an adaptor grammar~\cite{} based on $\mathcal{G}$ and some
parameters $\theta$.

>>>>

\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{amsmath} 
\usepackage{algpseudocode}
\usepackage{ amssymb }
\usepackage{listings}
\usepackage{syntax}
\usepackage{mdframed}
\usepackage{graphicx}

@@

Our probabilistic model is motivated by the fact that while $\mathcal{G}$
represents the language in which solutions are written, in any particular
domain there is a distribution $\mathcal{H}$ over solutions that is itself drawn
according to an adaptor grammar prior~\cite{} based on $\mathcal{G}$ and some
parameters $\theta$. $\mathcal{H}$ is a probabilistic context-free grammar
(PCFG), and this PCFG generates programs $\{e_n\}$ that solve problems in the
domain. Finally, any candidate solution $s_n$ is scored according to its
likelihood given the latent $e_i$. Thus we have the generative model pictured in
Figure~\ref{fig:generativeModel}.

\begin{figure}[h] 
  \includegraphics[width=.3\linewidth]{./figures/basicGM.pdf}  
  \caption{Generative model for expressions.\label{fig:generativeModel}}  
\end{figure}

The likelihood function $p_{s|e}(\cdot | e)$ takes the place in our framework that the fitness function would take in genetic programming.

How does the specification of the generative model relate to the specification
of the problems our agent needs to solve? The idea is that for any objective
function $ r:S \rightarrow \mathcal{D}$, where $S$ is the set of solutions and
$\mathcal{D}$ is the domain, we can map $\mathcal{D}$ to the interval $[0,1]$,
thereby generating a likelihood function that scores candidate solutions
relative to the latent program $e_n$.

How can we learn in this model? Our goal is to learn the distribution
$\mathcal{H}$.

\section{Potential application}
\begin{itemize}
  \item Learning GP kernels
  \item Learning control circuits (signals \& systems book)
  \item Gen audio
    Ryan suggested this. MAX MSP and its open source version (Pd or
    Pd-extended) allow composers to generate sounds by using a graphical
    programming language. 
  \item Gen vis
  \item Compositional object detectors
  \item Constructive solid geometry
\end{itemize}

\subsection{Discovering the laws of physics}
One of the main challenges that scientists face when discovering theories that
explain the world's phenomena is that they must perform inference over the
latent properties of objects in the world and the way in which these properties
relate to each other and the observable world. For example, Newton's first law
of motion postulates that things called forces exist in the world, and that
they obey the property that their vector sum is equal to the mass times the
acceleration of the object on which they act.

Here we consider simpler example of this joint inference problem: suppose you
observe various sets of charged particles interacting, but you do not know that
they are charged or that the property of electromagnetic charge exists. How can
you learn both that each particle has a charge, which is a positive or negative
scalar property, and that Coulomb's law dictates the force between those
particles.

Understanding Coulomb's law consists of holding that the following are true: 1)
every particle has a unique charge and, under normal circumstances, the charge
of a particle does not change. 2) Charge is a scalar quantity on the real line
that can take values on in the reals. 3) The force on any object $i$ with charge $q_i$ at location $r_i$ due to charged object j with charge $q_j$ at location $r_j$ is

\begin{align}
  F_i \propto \frac{q_i q_j}{|r_i - r_j|^2}
\end{align}

To simplify matters, and remove the complexities of vector arithmetic, we can
suppose that this is a one-dimensional problem:

\begin{align}
  F_i \propto \frac{q_i q_j}{(r_i - r_j)^2} \text{sign}(r_i - r_j)
\end{align}

Conceptually, this force law has two components. The first is that the force is
inversely propotional to the distance between the objects squared. The second
is that the force is proportional to the product of the charges. The scientist
can learn these two facts separately. By comparing the force between two
objects at different distances, she can discover the inverse squared distance
behavior. By noting that the force differs by the same multiplicative factor
when you swap object a for object c regardless of the paired object b.

In order to do this though, the scientist needs to be able to focus her
attention an different scenarios.
